---
title: "Data_processing"
output: html_document
date: "2026-08-28"
---



# --------------- Process sort-seq data

# Load DimSum output table
```{r}
# Download the Dimsum output table from FigShare and load into a datatable named LM:
  load(file.path(figshare_dir, "dimsum_local_maxima_2024_variant_data_merge.RData"))  
  LM<-as.data.table(variant_data_merge)
  LM$nt_seq<-toupper(LM$nt_seq)
  LM<-LM[, !c("WT", "STOP", "STOP_readthrough", "barcode_valid", "constant_region", "permitted", "too_many_substitutions", "mixed_substitutions")] ## Get rid of useless columns
  
# Set new column names, to ease the analysis downstream.  
  newnames<-c() 
  for(k in 1:length(colnames(LM))){
    newnames<-c(newnames, unlist(strsplit(colnames(LM)[k], "_e1"))[1])}
  colnames(LM)<-newnames
  
# Create 18 subdatatables, each with the data of one replicate and one drug:
  LM_r1<-LM[, .SD, .SDcols = names(LM) %like% "r1|nt_seq|aa_seq|WT"]
  newnames<-c() ;for(k in 1:length(colnames(LM_r1))){
  newnames<-c(newnames, unlist(strsplit(colnames(LM_r1)[k], "0"))[2])}; colnames(LM_r1)<-newnames
  
  LM_r2<-LM[, .SD, .SDcols = names(LM) %like% "r2|nt_seq|aa_seq|WT"]
  newnames<-c() ;for(k in 1:length(colnames(LM_r2))){
  newnames<-c(newnames, unlist(strsplit(colnames(LM_r2)[k], "0"))[2])}; colnames(LM_r2)<-newnames
  
  LM_r3<-LM[, .SD, .SDcols = names(LM) %like% "r3|nt_seq|aa_seq|WT"]
  newnames<-c() ;for(k in 1:length(colnames(LM_r3))){
  newnames<-c(newnames, unlist(strsplit(colnames(LM_r3)[k], "0"))[2])}; colnames(LM_r3)<-newnames
  
# Create the columns 'reads_allbins' (read count for each variant across bins), 'treatment' and 'replicate':
  LM_r1[, reads_allbins:=sum(c(r1p4,r1p5,r1p6,r1p7,r1p8,r1p9,r1p10)), by=1:nrow(LM_r1)][, replicate:=1]
  LM_r2[, reads_allbins:=sum(c(r2p4,r2p5,r2p6,r2p7,r2p8,r2p9,r2p10)), by=1:nrow(LM_r2)][, replicate:=2]
  LM_r3[, reads_allbins:=sum(c(r3p4,r3p5,r3p6,r3p7,r3p8,r3p9,r3p10)), by=1:nrow(LM_r3)][, replicate:=3]

```

#  Calculate the normalised readthrough values by:
a) Divide the number of reads of each variant by the total number of reads of that bin.
b) Normalise by the percentage of that bin in the total population (data in the Influx PDFs)
c) Finally, multiply by the mean mCherry value of each bin (stored in the 'gates_boundaries_LM' file.)
```{r}
# The mean mCherry values of each gate for each drug are stored in 'gates_boundaries' dtbl Load it: 
  
gates_boundaries_LM <- data.table(
  Replicate = rep(1:3, each = 8),
  Gate      = rep(paste0("P", 4:11), times = 3),
  Mean_value = c(
      3.57,  9.79, 18.99, 35.52, 70.47, 152.03, 314.84, 2878.00,  # Replicate 1
      3.69,  9.77, 19.03, 35.41, 70.42, 152.76, 319.76, 2828.43,  # Replicate 2
      3.65,  9.79, 19.04, 35.64, 70.66, 151.42, 315.33, 2974.01   # Replicate 3
  )
)

  r1_bins<-gates_boundaries_LM[Replicate==1, round(((Mean_value)/2878)*100,2)]
  r2_bins<-gates_boundaries_LM[Replicate==2, round(((Mean_value)/2828)*100,2)]
  r3_bins<-gates_boundaries_LM[Replicate==3, round(((Mean_value)/2974)*100,2)]
  
LM_r1[, c("r1p4_norm", "r1p5_norm", "r1p6_norm", "r1p7_norm", "r1p8_norm", "r1p9_norm", "r1p10_norm", "r1p11_norm"):=.((r1p4/(sum(r1p4)))*0.2791, (r1p5/(sum(r1p5)))*0.1249, (r1p6/(sum(r1p6)))*0.1613, (r1p7/(sum(r1p7)))*0.1498, (r1p8/(sum(r1p8)))*0.0809, (r1p9/(sum(r1p9)))*0.0364, (r1p10/(sum(r1p10)))*0.0112, (r1p11/(sum(r1p11)))*0.0125)][, sum_normreads:=sum(c(r1p4_norm, r1p5_norm, r1p6_norm, r1p7_norm, r1p8_norm, r1p9_norm, r1p10_norm)), by=1:nrow(LM_r1)][, fitness_rep:=((r1p4_norm*r1_bins[1] + r1p5_norm*r1_bins[2] + r1p6_norm*r1_bins[3] + r1p7_norm*r1_bins[4] + r1p8_norm*r1_bins[5] + r1p9_norm*r1_bins[6] + r1p10_norm*r1_bins[7])/sum_normreads), by=1:nrow(LM_r1)]

LM_r2[, c("r2p4_norm", "r2p5_norm", "r2p6_norm", "r2p7_norm", "r2p8_norm", "r2p9_norm", "r2p10_norm", "r2p11_norm"):=.((r2p4/(sum(r2p4)))*0.2791, (r2p5/(sum(r2p5)))*0.1249, (r2p6/(sum(r2p6)))*0.1613, (r2p7/(sum(r2p7)))*0.1498, (r2p8/(sum(r2p8)))*0.0809, (r2p9/(sum(r2p9)))*0.0364, (r2p10/(sum(r2p10)))*0.0112, (r2p11/(sum(r2p11)))*0.0125)][, sum_normreads:=sum(c(r2p4_norm, r2p5_norm, r2p6_norm, r2p7_norm, r2p8_norm, r2p9_norm, r2p10_norm)), by=1:nrow(LM_r2)][, fitness_rep:=((r2p4_norm*r2_bins[1] + r2p5_norm*r2_bins[2] + r2p6_norm*r2_bins[3] + r2p7_norm*r2_bins[4] + r2p8_norm*r2_bins[5] + r2p9_norm*r2_bins[6] + r2p10_norm*r2_bins[7])/sum_normreads), by=1:nrow(LM_r2)]

LM_r3[, c("r3p4_norm", "r3p5_norm", "r3p6_norm", "r3p7_norm", "r3p8_norm", "r3p9_norm", "r3p10_norm", "r3p11_norm"):=.((r3p4/(sum(r3p4)))*0.2791, (r3p5/(sum(r3p5)))*0.1249, (r3p6/(sum(r3p6)))*0.1613, (r3p7/(sum(r3p7)))*0.1498, (r3p8/(sum(r3p8)))*0.0809, (r3p9/(sum(r3p9)))*0.0364, (r3p10/(sum(r3p10)))*0.0112, (r3p11/(sum(r3p11)))*0.0125)][, sum_normreads:=sum(c(r3p4_norm, r3p5_norm, r3p6_norm, r3p7_norm, r3p8_norm, r3p9_norm, r3p10_norm)), by=1:nrow(LM_r3)][, fitness_rep:=((r3p4_norm*r3_bins[1] + r3p5_norm*r3_bins[2] + r3p6_norm*r3_bins[3] + r3p7_norm*r3_bins[4] + r3p8_norm*r3_bins[5] + r3p9_norm*r3_bins[6] + r3p10_norm*r3_bins[7])/sum_normreads), by=1:nrow(LM_r3)]

```

# Merge the 3 replicate dtbls into one, named 'LM_dtbl'.
```{r}
# Remove the 'r1' or 'r2' of the column names, to be able to merge all the datatables in one:
  cnames<-c("nt_seq", "aa_seq", "p4", "p5", "p6", "p7", "p8", "p9", "p10", "p11", "reads_allbins", "replicate", "p4_norm", "p5_norm", "p6_norm", "p7_norm", "p8_norm", "p9_norm",  "p10_norm",  "p11_norm", "sum_normreads", "fitness_rep")

  colnames(LM_r1)<-cnames; colnames(LM_r2)<-cnames; colnames(LM_r3)<-cnames
  
# Merge the two replicates of each treatment in one datatable:
  LM_dtbl<-rbind(LM_r1, LM_r2, LM_r3)
  
# Set fitness_rep to NA for low coverage (<10) variants, so they are not used for RT calculation:    
  
  LM_dtbl[replicate==1 & reads_allbins<10, .N] # n=47
  LM_dtbl[replicate==2 & reads_allbins<10, .N] # n=51
  LM_dtbl[replicate==3 & reads_allbins<10, .N] # n=49
  LM_dtbl[reads_allbins<10, fitness_rep:=NA]
  
# Set the replicate column as factor, instead of numeric:
  LM_dtbl$replicate<-as.factor(LM_dtbl$replicate)

  
# Calculate the mean readthrough efficiency and the sdeviation ('RT' and 'sd_RT' columns).
  LM_dtbl[, c("RT", "sd_RT"):=.(mean(fitness_rep, na.rm=T), sd(fitness_rep, na.rm=T)), by=nt_seq]

  
# Create the 'reads_allbins_r1', 'reads_allbins_r2' and 'reads_allbins_median' columns: 
  for (i in c(2,3)){
  LM_dtbl[replicate==1, reads_allbins_r1:=reads_allbins, by="nt_seq"]
  LM_dtbl[replicate==i, reads_allbins_r1:=LM_dtbl[nt_seq==nt_seq & replicate==1, reads_allbins]]
  }
  
  for (i in c(1,3)){
  LM_dtbl[replicate==2, reads_allbins_r2:=reads_allbins, by="nt_seq"]
  LM_dtbl[replicate==i, reads_allbins_r2:=LM_dtbl[nt_seq==nt_seq & replicate==2, reads_allbins]]
  }

  for (i in c(1,2)){
  LM_dtbl[replicate==3, reads_allbins_r3:=reads_allbins, by="nt_seq"]
  LM_dtbl[replicate==i, reads_allbins_r3:=LM_dtbl[nt_seq==nt_seq & replicate==3, reads_allbins]]
  }

# the column 'reads' shows the median of reads_allbins across replicates:
  LM_dtbl[, reads:=round(matrixStats::rowMedians(as.matrix(.SD), na.rm=T), 0), .SDcols=c("reads_allbins_r1", "reads_allbins_r2", "reads_allbins_r3")]
  
```

# Create the RT_binomial column (needed for model fitting later on):
```{r, warning=F}
  LM_dtbl[, RT_binomial:=RT/100, by=.I]
```


# Check the distribution of the non-stop TP53 control variant across bins. Ideally, it should be exclusively found in P11:
```{r, warning=F, fig.width=4.5, fig.height=5}

# Calculate the percentage of P11 reads taken up by the control variant:
  (LM_dtbl[nt_seq==NONSTOP_CTRL_SEQ & 
               replicate==1, p11]/LM_dtbl[replicate==1, sum(p11)])*100 # -> 32%
  
  
  nonstop_ctrl<-LM_dtbl[nt_seq==NONSTOP_CTRL_SEQ &
                    replicate==1, c("nt_seq","p4","p5","p6","p7","p8", "p9","p10", "p11","p4_norm",
                                    "p5_norm","p6_norm","p7_norm","p8_norm", "p9_norm", "p10_norm", "p11_norm", "replicate")]
  nonstop_ctrl_long <- as.data.table(gather(nonstop_ctrl, bin, norm_reads, p4:p11_norm, factor_key=TRUE))
    
    p1<-ggplot(nonstop_ctrl_long[bin%chin%c("p4","p5","p6","p7","p8", "p9", "p10", "p11")], aes(x=bin, y=norm_reads)) + 
      geom_bar(stat="identity", fill="#FF6666") + p + 
      ylab("Read counts (non-stop variant)") + theme(axis.title.x = element_blank())
    
    p2<-ggplot(nonstop_ctrl_long[bin%chin%c("p4_norm","p5_norm","p6_norm","p7_norm","p8_norm","p10_norm", "p11_norm")], 
               aes(x=bin, y=norm_reads)) + geom_bar(stat="identity", fill="#FF6666") + p
    
  p2; p1
```


```{r, warning=F}
pdf(file.path(plot_dir, "Control_variant.pdf"), height = 5, width = 5)
print(p1); dev.off()
```



# Remove the nonstop_ctrl variant + 'p11' and 'p11_norm' columns from LM_dtbl, to avoid confusions in downstream analyses:
- 'p11' and 'p11_norm' columns are not used to compute RT for other variants, so pointless to keep them
- we just wanted to make sure it was mostly found in P11. Now, we don't need it anymore.

```{r, warning=F}
  LM_dtbl<-LM_dtbl[nt_seq!=NONSTOP_CTRL_SEQ]
  LM_dtbl[, p11:=NULL]
  LM_dtbl[, p11_norm:=NULL]
```

# Remove a variant suspucious of cryptic splicing:
```{r, warning=F}
  LM_dtbl<-LM_dtbl[nt_seq!=CRYPTIC_SPLICING_SEQ]
```

# Show the distribution of normalised readcounts across bins for some randomly sampled variants (n=10): 
```{r, warning=F, fig.height=3, fig.width=12}

  LMsubset<-LM_dtbl[replicate==1, c("p4_norm", "p5_norm", "p6_norm", "p7_norm", "p8_norm", "p9_norm", "p10_norm")]
  setnames(LMsubset, c("p4_norm", "p5_norm", "p6_norm", "p7_norm", "p8_norm", "p9_norm", "p10_norm"), c("p4", "p5", "p6", "p7", "p8", "p9", "p10"))
  sampled<-sample(c(1:5208), 10)
  level_order <- c("p4", "p5", "p6", "p7", "p8", "p9", "p10") ## This is just to sort the x axis of GGplot.
  
  j<-1; plot_list = list()
  for (k in sampled){
    x<-data.frame(x1=colnames(LMsubset), normalised_counts=as.numeric(LMsubset[k,]))
    r<-ggplot(x, aes(x=factor(x1, level=level_order), y=normalised_counts)) + geom_bar(stat="identity", fill="blue3") + xlab("Sorting bins")
    plot_list[[j]]=r
    j<-j+1
  }
  
  p2<-ggarrange(plot_list[[1]],plot_list[[2]],plot_list[[3]],plot_list[[4]],plot_list[[9]],
            plot_list[[5]],plot_list[[6]],plot_list[[7]],plot_list[[8]],plot_list[[10]],nrow=2,ncol=5)
  p2
  
```

```{r, warning=F}
pdf(file.path(plot_dir, "Bins_distribution_ten_variants.pdf"), height = 3, width = 12)
print(p2); dev.off()
```
# Associate each variant DimSum info to the mutagenesis info (by merging LM_dtbl with each of the 14 dataframes containing a different sublibrary/mutagenesis_strategy):
```{r, warning=F}

# Read the 14 dtbls:
  down_blocks_combinations<-readRDS(file = file.path(figshare_dir, "down_blocks_combinations.rds"))
  up_blocks_combinations<-readRDS(file = file.path(figshare_dir, "up_blocks_combinations.rds"))
  tetramers<-readRDS(file = file.path(figshare_dir, "tetramers.rds"))
  df_singles<-readRDS(file = file.path(figshare_dir, "df_singles.rds"))
  df_doubles<-readRDS(file = file.path(figshare_dir, "df_doubles.rds"))
  trimers<-readRDS(file = file.path(figshare_dir, "trimers.rds"))
  df_doubles2<-readRDS(file = file.path(figshare_dir, "df_doubles2.rds"))
  aa_AQP4_final<-readRDS(file = file.path(figshare_dir, "aa_AQP4_final.rds"))
  aa_MAPK10_final<-readRDS(file = file.path(figshare_dir, "aa_MAPK10_final.rds"))
  aa_OPRK1_final<-readRDS(file = file.path(figshare_dir, "aa_OPRK1_final.rds"))
  aa_OPRL1_final<-readRDS(file = file.path(figshare_dir, "aa_OPRL1_final.rds"))
  single_proline<-readRDS(file = file.path(figshare_dir, "single_proline.rds"))
  double_proline<-readRDS(file = file.path(figshare_dir, "double_proline.rds"))
  triple_proline<-readRDS(file = file.path(figshare_dir, "triple_proline.rds"))
  only_upstream<-readRDS(file = file.path(figshare_dir, "only_upstream.rds"))
  only_downstream<-readRDS(file = file.path(figshare_dir, "only_downstream.rds"))
  
# Store this 14 dataframes in input_list
  input_list<-list(down_blocks_combinations, up_blocks_combinations, tetramers, df_singles, df_doubles, trimers, df_doubles2, aa_AQP4_final, 
                   aa_MAPK10_final, aa_OPRK1_final, aa_OPRL1_final, single_proline, double_proline, triple_proline, only_upstream, only_downstream)
# Merge each of the 14df with LM_dtbl and store them separately in output_list.
    output_list<-list()
    for (k in 1:length(input_list)){
    r<-inner_join(input_list[[k]], LM_dtbl, by="nt_seq")
    output_list[[k]]=as.data.table(r)}
    
# Merge the 14 dataframes from output_list together to generate LM_dt, which contains info of all variants across all mutagenesis types: some sequences are duplicated, because they belong to two or more sublibraries and they are represetned with one row for each mutagenesis type.That's why the nrow of LM_dt is 5231 whereas the number of unique variants is 5071.
    LM_dtbl<-output_list[[1]]
    for  (i in 2:length(output_list)){
    LM_dtbl<-full_join(LM_dtbl, output_list[[i]], by=NULL)}    
    
  # There is a misannotation where some variants classified as "double_mutants_11-15nts_downstream" are instead single mutants. Get rid of these using hamming distance:
    LM_dtbl[Gene=="AQP4" & Mutation_type=="double_mutants_11-15nts_downstream", 
            isTRUEdoubmut:=as.vector(stringdist::stringdist(nt_seq, "AGAGTATTGTCTTCAGTATGACTAGAAGATCGCACT", method="hamming"))>1] # AQP4
    LM_dtbl[Gene=="OPRK1" & Mutation_type=="double_mutants_11-15nts_downstream", 
            isTRUEdoubmut:=as.vector(stringdist::stringdist(nt_seq, "GACATCGATGGGATGAATAAACCAGTATGACTAGTCGTGGAGATG", method="hamming"))>1] # OPRK1
    LM_dtbl[Gene=="MAPK10" & Mutation_type=="double_mutants_11-15nts_downstream", 
            isTRUEdoubmut:=as.vector(stringdist::stringdist(nt_seq, "CCCCTGGGTTGTTGCAGGTGACTAGCCGCCTGCCTGCGAAACCCAGCG", method="hamming"))>1] # MAPK10
    LM_dtbl[Gene=="OPRL1" & Mutation_type=="double_mutants_11-15nts_downstream", 
            isTRUEdoubmut:=as.vector(stringdist::stringdist(nt_seq, "ACGGTACCGCGGCCCGCATGACTAGGCGTGGACCTG", method="hamming"))>1] # OPRL1

    LM_dtbl<-LM_dtbl[isTRUEdoubmut==T | is.na(isTRUEdoubmut), ]
```

# --- Do some arrangements. Basically change a few column names and duplicate variants which belong to more than one mutagenesis category:

#Single nt variants have their mutagenesis data in columns "Position_mutated" and "Mutation". With this script, I move the informatio to "Position_mutated_1" and "Mutation_1" columns, where the double_nt variants have their informatioon (is more practical for downstream analysis).
But don't elimiante "Position_mutated" and "Mutation" because they still have mutagenesis information in other variants.
```{r, warning=F}
    LM_dtbl<-LM_dtbl[Mutation_type=="Single_mutants_downstream_allnts", c("Position_mutated_1", "Mutation_1") :=.(Position_mutated, Mutation)]
    LM_dtbl<-LM_dtbl[Mutation_type=="Single_mutants_downstream_allnts", c("Position_mutated", "Mutation") :=.(NA, NA)]
```

# Label all the aa_doublemutants variants as such ('remove the gene name'):
```{r, warning=F}   
LM_dtbl[Mutation_type%chin%c("aa_doublemutants_AQP4", "aa_doublemutants_MAPK10", "aa_triplemutants_OPRK1", "aa_triplemutants_OPRL1"),
        Mutation_type:="aa_double_triple_mutants"]
```

# Create duplicated entries for the 'CTAG_randomized variants' that belong also to the 'double_mutants_downstream_+1/+10nts' group. Right now they are only labelled as 'CTAG_randomized variants' and when we select the 'double_mutants_downstream_+1/+10nts' they are missing.
```{r, warning=F}   
  test<-LM_dtbl[Mutation_type=="CTAG_randomization"]
# Select those with Hamming Distance of 2, which are those belonging to the 'double_mutants_downstream_+1/+10nts' group:
  test[, double_mut_TRUE:= as.vector(stringdist::stringdist(Mutation, "CTAG", method="hamming"))==2, by=1:nrow(test)]
  test<-test[double_mut_TRUE==T]
  test[, WT_seq:="CTAG"]
  list<-mapply(function(x, y) which(x != y), strsplit(test$Mutation, ""), strsplit(test$WT_seq, ""))
    
  for (i in 1:nrow(test)){
    test$Position_mutated_1[i]<-list[1,i]
    test$Position_mutated_2[i]<-list[2,i]
    test$Mutation_1[i]<-s2c(test$Mutation[i])[test$Position_mutated_1[i]]
    test$Mutation_2[i]<-s2c(test$Mutation[i])[test$Position_mutated_2[i]]
    }
  test[, Mutation_type:="double_mutants_downstream_+1/+10nts"]
  test[, WT_seq:=NULL]; test[, double_mut_TRUE:=NULL]
  LM_dtbl<-rbind(LM_dtbl, test)
```


# Same but for the single mutants:
```{r, warning=F}   
  test<-LM_dtbl[Mutation_type=="CTAG_randomization"]
# Select those with Hamming Distance of 1, which are those belonging to the 'Single_mutants_downstream_allnts' group:
  test[, single_mut_TRUE:= as.vector(stringdist::stringdist(Mutation, "CTAG", method="hamming"))==1, by=1:nrow(test)]
  test<-test[single_mut_TRUE==T]; test[, WT_seq:="CTAG"]
  list<-mapply(function(x, y) which(x != y), strsplit(test$Mutation, ""), strsplit(test$WT_seq, ""))
    
  for (i in 1:nrow(test)){
    test$Position_mutated_1[i]<-list[i]
    test$Mutation_1[i]<-s2c(test$Mutation[i])[test$Position_mutated_1[i]]
    }
  test[, Mutation_type:="Single_mutants_downstream_allnts"]
  test[, WT_seq:=NULL]; test[, single_mut_TRUE:=NULL]
  LM_dtbl<-rbind(LM_dtbl, test)
```

# Label the chimera variants as such. They are now unlabelled (Gene=NA)
```{r, warning=F} 
  LM_dtbl[is.na(Gene), Gene:="Chimera"]
```

# After these rearengements, make sure we don't have unwanted duplications. We want a unique combination of 'nt_seq - Mutation_type - replicate' columns.
```{r, warning=F} 
  
  LM_dtbl[, .N] # 16548 rows

  # Print duplicated variants for that triple column combination:
    LM_dtbl[, .N, by = .(nt_seq, Mutation_type, replicate)][N>1]
  
  # Keep only one row per each combination:
    LM_dtbl <- unique(LM_dtbl, by = c("nt_seq", "Mutation_type", "replicate"))
    LM_dtbl[, .N] # 16425 rows
``` 

# Create the 'downstream_seq' column, which shows the sequence downstream of the stop codon. The column is set to NA for chimera and upstream variants:
```{r, warning=F, fig.height=5, fig.width=10}  
  
  # Gene-specific split patterns, which are the upstream_seq + stop codon:
  split_patterns <- c(
    AQP4   = "AGAGTATTGTCTTCAGTATGA",
    OPRK1  = "GACATCGATGGGATGAATAAACCAGTATGA",
    MAPK10 = "CCCCTGGGTTGTTGCAGGTGA",
    OPRL1  = "ACGGTACCGCGGCCCGCATGA"
  )

  LM_dtbl[ ,
    downstream_seq := mapply(
      function(seq, pat) strsplit(seq, pat, fixed = TRUE)[[1]][2],
      nt_seq,
      split_patterns[Gene]
    )
  ]
```

 
# Create the 'PTC' column -> Some mutations introduce a PTC, which are used as negative controls. The 'PTC' columns shows if the variant has a PTC ('yes') or not ('no'). 

```{r, warning=F, fig.height=5, fig.width=10} 

# The second PTC is found downstream:
  for (k in 1:nrow(LM_dtbl)){ 
    if (is.na(LM_dtbl$downstream_seq[k])){LM_dtbl$PTC[k]<-"no"} 
    else if (sum((substring(LM_dtbl$downstream_seq[k], seq(1, nchar(LM_dtbl$downstream_seq[k]), 3), 
                            pmin(seq(1, nchar(LM_dtbl$downstream_seq[k]), 3) + 3 - 1, 
                                 nchar(LM_dtbl$downstream_seq[k]))))%chin%c("TGA", "TAG", "TAA"), na.rm = TRUE)==0)
         {LM_dtbl$PTC[k]<-"no"} 
    else {LM_dtbl$PTC[k]<-"yes"}
  }

# The second PTC is found upstream:
  LM_dtbl[Mutation_type=="3nts_upstream_PTC_randomized" & Mutation%chin%c("TGA", "TAG", "TAA"), PTC:="yes"]
```

Plot:
```{r, warning=F, fig.height=4, fig.width=8} 
  p1<-ggplot(LM_dtbl[!duplicated(nt_seq) & Gene%chin%c("AQP4", "OPRK1", "MAPK10")], aes(x=PTC, y=RT)) + geom_boxplot(outliers = F) + 
    geom_jitter(alpha=.2) + facet_wrap(~Gene, scales="free") + p + ylab("Readthrough (%)") + xlab(">1 PTC")
  LM_dtbl[!duplicated(nt_seq), table(PTC)]
  
  p1
```

```{r, warning = FALSE}  
pdf(file.path(plot_dir, "PTC2x_controls.pdf"), height = 4, width = 8)
print(p1); dev.off()
```

# Create the WT column and factorize the gene column:
```{r, warning=F, fig.height=5, fig.width=10}

# Factorize gene column:
  LM_dtbl$Gene<-factor(LM_dtbl$Gene, levels=c("AQP4", "OPRK1", "MAPK10", "OPRL1", "Chimera"))

# WT column:
  LM_dtbl[, WT:="no"]
  LM_dtbl[nt_seq=="AGAGTATTGTCTTCAGTATGACTAGAAGATCGCACT", WT:="yes"] # AQP4
  LM_dtbl[nt_seq=="CCCCTGGGTTGTTGCAGGTGACTAGCCGCCTGCCTGCGAAACCCAGCG", WT:="yes"] # MAPK10
  LM_dtbl[nt_seq=="GACATCGATGGGATGAATAAACCAGTATGACTAGTCGTGGAGATG", WT:="yes"] # OPRK1
  LM_dtbl[nt_seq=="ACGGTACCGCGGCCCGCATGACTAGGCGTGGACCTG", WT:="yes"] # OPRL1

```

# Generate three new columns: 
- The RT_WTnorm column, where the RT of the WT is subtracted from each variant, by gene.
- pval_diff_from_WT, where we test the significance of the RT_WTnorm score
- The FDR_diff_from_WT. FDR correction of multitesting.
```{r, warning=F}

# RT_WTnorm column
  AQP4_wt_RT<-LM_dtbl[Gene=="AQP4" & WT=="yes", unique(RT)] 
  OPRK1_wt_RT<-LM_dtbl[Gene=="OPRK1" & WT=="yes", unique(RT)]
  MAPK10_wt_RT<-LM_dtbl[Gene=="MAPK10" & WT=="yes", unique(RT)]
  # Can't do it with OPRL1 bc we don't have the WT
  
  LM_dtbl[Gene=="AQP4", RT_WTnorm:=RT-AQP4_wt_RT]
  LM_dtbl[Gene=="OPRK1", RT_WTnorm:=RT-OPRK1_wt_RT]
  LM_dtbl[Gene=="MAPK10", RT_WTnorm:=RT-MAPK10_wt_RT]
  
# FDR_diff_from_WT column: does the variant have a RT significantly different from WT?
  
  # Get the RT values across replicates for the WTs and store in genes_WT:
    AQP4_wt<-LM_dtbl[Gene=="AQP4" & WT=="yes", unique(fitness_rep)] 
    OPRK1_wt<-LM_dtbl[Gene=="OPRK1" & WT=="yes", unique(fitness_rep)]
    MAPK10_wt<-LM_dtbl[Gene=="MAPK10" & WT=="yes", unique(fitness_rep)] 
    genes_WT<-rbind(AQP4_wt, OPRK1_wt, MAPK10_wt)
  
  # T.test variant vs WT across the 3 replicates, store in 'pval_diff_from_WT':
  genes<-c("AQP4", "OPRK1", "MAPK10")
  for (i in 1:length(genes)){
    LM_dtbl[Gene==genes[i], pval_diff_from_WT := {
      x <- fitness_rep[!is.na(fitness_rep)]
      if (length(x) < 2) NA_real_ else t.test(x, genes_WT[i,])$p.value
    }, by = nt_seq]
  }

# Adjust pvalue using BH/FDR:
  LM_dtbl[, FDR_diff_from_WT := p.adjust(pval_diff_from_WT, method = "BH")]

# How many genes have a significantly higher RT over WT, keeping the FDR at 0.05?
  LM_dtbl[!duplicated(nt_seq) & RT_WTnorm>0 & FDR_diff_from_WT<0.05, .N, by=Gene]
```



